 

### 2023-02-16 markdown/introduction.Rmd
quality control started

# Introduction

@@

@@

Further analysis is done with R and respective packages such as bioconductor.

>>>>

@@

#### Data processing

Further analysis can be done with R and respective packages such as bioconductor. Since there is no state of the art established, analysis varies upon experimental design.

After processing MaxQuant creates a directory containing all results as .txt file. The evidence.txt file include all peptide sequence matches (PSM) with their respective proteins and statistical parameters.

Example fo basic parameters and derivations include:

* Peptide sequence
* Mass to charge ratio (m/z) for all scans (eg. MS1, MS2)
	+ Mass
* Retention time
* Precursor Ion Fragment
	+ source of the detected ion also referred as mother ion
* Fraction of total spectrum 
* Base peak fraction
* Reporter intensity (RI)
	+ corrected RI
* Posterior error probability (PEP)

#### Object oriented programming

In order to streamline the analysis of multiple experiments, object oriented programming can be applied. The approach in R is to create a so called Q-feature object, which contains all variables and metadata in a hierarchical structure. The structure enables sub setting for further analysis.

#### Filtering

Peptides having a low abundance will be set to zero by the analysis. The value zero would imply that the sample does not contain the respective peptide, therefore replacing 0 with not applicable (NA) is crucial for understanding and interpretation of MS-data.

Peptide sequences matching to the reverse protein sequences (=decoy database) are considered as possible contaminants. These matches can be excluded from further analysis.

The author of the SCoPE2 \citep{Petelski2021} pipeline recommends calculating the sample to carrier ration (SCR). This can be done by dviding the ion intensity of the sample by the reporter ion intensity of the carrier channel. However in our experiment no carrier and reference channel is used.

The next step for quality control is the exclusion of samples with a high false discovery rate (FDR). When applying multiple statistical testing (e.g. t-Test) the obtained p-values can be considered as biased, because the probability to observe a significant will iteratively increase with each test performed. Corrections in statistics are an approach to compensate for the multiplicity of testing. There are many ways to do this compensation like the Bonferroni method or Benjamini-Hochberg`s FDR. In Mass spectrometry the common “way to go” is calculating a false discovery rate, by dividing false PSMs (=hit of the decoy database) through the total number of PSMs above the peptide-spectrum matching score. The peptides pectrum matching score is defined as -10log10(p). Whereas the p-value is defined that the hit is done by chance. The calculation of the score is highly dependent on the data acquisition method used. MaxQuant uses Andromeda , an integrated search engine. Proteome Discoverer from Thermo Fisher utilizes different engines such as Mascot or Minora. As published by J.Cox in 2011 Mascot and Andromeda showed similar performance when comparing FDR values as a function of coverage. However the observed performance can be lower when dealing with a decreased coverage \citep{Cox2011}.The threshold for accepting an FDR of an individual PSM is described as q-value.

Further filter steps can include correction between multiple runs. This kind of process need the addition of a reference channel. However when observing a single run, these steps are not crucial for the upcoming analysis.

@@
